---
title: "hw1"
author: "Yanhao Shen"
output: html_document
date: "2024-09-18"
---

Q1.
```{r}
data('penguins', package = 'palmerpenguins')
summary(penguins)
```
The dataset is about penguins, and it has 8 variables: species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex, and year. The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns. The mean flipper length is `r mean(penguins$flipper_length_mm, na.rm=TRUE)`.
```{r}
library(ggplot2)
scatter_plot <- ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm, color=species)) + geom_point()
scatter_plot
ggsave('scatter_plot.pdf', height = 4, width =6)
```

Q2.
```{r}
df <- data.frame(random_sample = rnorm(10),
                 logical_vector = rnorm(10) > 0,
                 character_vector = sample(letters, 10),
                 factor_vector = factor(sample(c('Low',
                                                 'Medium',
                                                 'High'),
                                               10, replace = TRUE))
                 )
df
```
```{r}
mean(df$random_sample) #works because it is numeric
mean(df$logical_vector) #works because TRUE are treated as 1, FALSE as 0
mean(df$character_vector) # doesn’t work because character are not numeric
mean(df$factor_vector) #doesn’t work because factors are not numeric
```
```{r eval=FALSE}
as.numeric(df$logical_vector) #True will be 1, and False will be 0
as.numeric(df$character_vector) #nothing help
as.numeric(df$factor_vector) #factors will be converted to underlying numbers.
```
Numeric and Logical variables can have a mean because R can interpret them as numbers. Character and factors can’t have a mean without conversion. While after conversion, characters will be NA, and factors will be their underlying numbers. in this df, 3 as Medium, 2 as Low, and 1 as High.
